# Solution

## Task 1

(a) 

```{r, eval = TRUE}
# taste vs lactic acid
plot(Taste ~ Lactic.Acid, data = cheese, xlab = "Lactic acid concentration", ylab = "Taste score")

# taste vs H2S
plot(Taste ~ H2S, data = cheese, xlab = "H2S concentration", ylab = "Taste score")
```

(b) 

```{r}
# taste vs log(lactic acid)
plot(Taste ~ log(Lactic.Acid), data = cheese, xlab = "Log lactic acid concentration", ylab = "Taste score")

# taste vs log(H2S)
plot(Taste ~ log(H2S), data = cheese, xlab = "Log H2S concentration", ylab = "Taste score")
```

(c) Based on the above four plots, it is clear that log(**H2S**) is preferred over **H2S** in explaining **Taste**. The difference between **Lactic Acid** and log(**Lactic Acid**)) is little and therefore we will select the original variable **Lactic Acid** as the predictor. 

To build a multiple linear regression between **Taste** (as the response) and **Lactic Acid** and log(**H2S**) (as the predictors), we use
```{r}
model <- lm(Taste ~ Lactic.Acid + log(H2S), data=cheese)
summary(model)
```

When **Lactic Acid** and log(**H2S**) are both zero, the mean cheese taste score is $-27.59$. The mean cheese taste score increases by $19.89$ when **Lactic.Acid** increases by one unit and *all other predictors are held constant*. The mean cheese taste score increases by $3.95$ when log(**H2S**) increases by one unit and *all other predictors are held constant*. 

(d)

```{r, echo=FALSE}
# calculate according to the formula
X <- model.matrix(~ Lactic.Acid + log(H2S), data=cheese)
Y <- cheese$Taste

XtX <- t(X) %*% X
XtY <- t(X) %*% Y

beta.hat <- solve(XtX) %*% XtY
beta.hat

#from R output
coef(model)
```

(e) 
```{r}
# calculate according to the formula
YtY <- t(Y) %*% Y
RSS <- YtY - t(XtY) %*% beta.hat
sigma2_hat <- RSS/(nrow(cheese)-3)
sigma2_hat

# from R output
summary(model)$sigma
```


